---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README_files/README-",
  fig.width = 6,
  fig.height = 6
)
```

# cranly


[**cranly**](https://github.com/ikosmidis/cranly) provides core visualisations and summaries for the CRAN package database. It is aimed mainly as an analytics tool for developers to keep track of their CRAN packages and profiles, as well as those of others, which, at least for me, is proving harder and harder with the rapid growth of the [CRAN](https://cran.r-project.org) ecosystem.


The package provides comprehensive methods for cleaning up and organising the information in the CRAN package database, for building package directives networks (depends, imports, suggests, enhances) and collaboration networks, and for computing summaries and producing interactive visualisations from the resulting networks. Network visualisation is through the [**visNetwork**](https://CRAN.R-project.org/package=visNetwork) package. The package also provides functions to coerce the networks to igraph <https://CRAN.R-project.org/package=igraph> objects for further analyses and modelling.

### Installation

Install the development version from github:
``` r
# install.packages("devtools")
devtools::install_github("ikosmidis/cranly")
```

### Collaboration and package directives networks in CRAN

The first step in the **cranly** workflow is to try and
"clean-up" the package and author names in the data frame that results
from a call to `tools::CRAN_package_db()`
```{r}
p_db <- tools::CRAN_package_db()
package_db <- clean_CRAN_db(p_db)
```
The CRAN database we use is from
```{r}
attr(package_db, "timestamp")
```

#### Package directives networks

The package directives network can then be built using
```{r}
package_network <- build_network(package_db)
```
`package_network` can then be interrogated using extractor methods
(see, `?package_by`). For example, my packages can be extracted as follows
```{r}
my_packages <- package_by(package_network, "Ioannis Kosmidis")
my_packages
```
and their sub-network of directives can be summarized in an interactive visualization, a shapshot of which is below
```{r}
visualize(package_network, package = my_packages, title = TRUE, legend = TRUE)
```

We can also compute package summaries and plot "Top-n" lists according to the various summaries
```{r}
package_summaries <- summary(package_network)
plot(package_summaries, according_to = "n_imported_by", top = 20)
plot(package_summaries, according_to = "page_rank", top = 20)
```

#### Collaboration networks
The collaboration network can also be built using a similar call
```{r}
author_network <- build_network(package_db, perspective = "author")
```
and the extractor functions work exactly as they did for the package directives network. For example, my collaboration network results can be summarized as an interactive visualization, a shapshot of which is below
```{r}
visualize(author_network, author = "Ioannis Kosmidis")
```

"Top-n" collaborators according to various summaries can again be computed
```{r}
author_summaries <- summary(author_network)
plot(author_summaries, according_to = "n_collaborators", top = 20)
plot(author_summaries, according_to = "n_packages", top = 20)
plot(author_summaries, according_to = "page_rank", top = 20)
```

Well, the usual suspects...

Check the package vignettes for a more comprehensive tour of the package and for network visualisations on authors with orders of magnitude larger collaboration networks than mine.
