---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README_files/README-",
  fig.width = 6,
  fig.height = 6
)
```

# cranly

[**cranly**](https://github.com/ikosmidis/cranly) provides core visualisations and summaries for the CRAN package database. The package provides comprehensive methods for cleaning up and organising the information in the CRAN package database, for building package directives networks (depends, imports, suggests, enhances) and collaboration networks, and for computing summaries and producing interactive visualisations from the resulting networks. Network visualisation is through the [**visNetwork**](https://CRAN.R-project.org/package=visNetwork) package. The package also provides functions to coerce the networks to igraph <https://CRAN.R-project.org/package=igraph> objects for further analyses and modelling.

### Installation

Install the development version from github:
``` r
# install.packages("devtools")
devtools::install_github("ikosmidis/cranly")
```

### Collaboration and package directives networks in CRAN

The first step in the workflow **cranly** sets is to try and
"clean-up" the package and author names in the data frame that results
from a call to `tools::CRAN_package_db()`
```{r}
p_db <- tools::CRAN_package_db()
package_db <- clean_CRAN_db(p_db)
```

#### Package directives networks

The package directives network can then be built using
```{r}
package_network <- build_network(package_db)
```
`package_network` can then be interrogated using extractor methods
(see, `?package_by`). For example, my packages can be extracted as follows
```{r}
my_packages <- package_by(package_network, "Ioannis Kosmidis")
my_packages
```
and their sub-network of directives can be summarized in an interactive visualization, a shapshot of which is below
```{r}
visualize(package_network, package = my_packages, title = TRUE, legend = TRUE)
```

We can also compute package summaries and plot "Top-n" lists according to the various summaries
```{r}
package_summaries <- summary(package_network)
plot(package_summaries, according_to = "n_imported_by", top = 20)
plot(package_summaries, according_to = "page_rank", top = 20)
```

#### Collaboration networks
The collaboration network can also be built using a similar call
```{r}
author_network <- build_network(package_db, perspective = "author")
```
and the extractor functions work exactly as they did for the package directives network. For example, [Brian Ripley's](https://en.wikipedia.org/wiki/Brian_D._Ripley) massive collaboration network results in an nice interactive visualization, a shapshot of which is below
```{r}
visualize(author_network, author = "Brian Ripley")
```

"Top-n" collaborators according to various summaries can again be computed
```{r}
author_summaries <- summary(author_network)
plot(author_summaries, according_to = "n_collaborators", top = 20)
plot(author_summaries, according_to = "n_packages", top = 20)
plot(author_summaries, according_to = "page_rank", top = 20)
```

Well, the usual suspects... Dynamic networks are already in the wishlist
