---
title: "CRAN top 10s and interactive package and collaboration networks"
author: "[Ioannis Kosmidis](http://www.ucl.ac.uk/~ucakiko/)"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CRAN top 10s and interactive package and collaboration networks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Preparing today's CRAN package database

Donwload today's CRAN database and clean and organise author names, depends, imports, suggests, enhances
```{r}
library("cranly")
packages_db <- clean_CRAN_db()
```

The resulting dataset carries the timestamp of when it was put together
```{r}
attr(packages_db, "timestamp")
```

# Network of CRAN packages

Extract edges and nodes for the CRAN package network
```{r}
package_network <- build_cranly_network(object = packages_db)
```

Compute statistics for the package network
```{r}
## Global package network statistics
package_summaries <- summary(package_network)
```

Top 10 packages according to various statistics
```{r}
library("dplyr")
library("ggplot2")
## package_summaries %>%
##     arrange(eigen_centrality) %>%
##     mutate(package = factor(package, ordered = TRUE, levels = package)) %>% tail(50) %>%
##     ggplot(aes(y = eigen_centrality, x = package)) + geom_bar(stat = "identity") + coord_flip()
package_summaries %>% arrange(desc(n_imported_by)) %>%
    select(package, n_authors, n_imports, n_imported_by) %>% head(10)
package_summaries %>% arrange(desc(page_rank)) %>%
    select(package, n_authors, n_imports, n_imported_by, page_rank) %>% head(10)
package_summaries %>% arrange(desc(betweenness)) %>%
    select(package, n_authors, n_imports, n_imported_by, betweenness) %>% head(10)
package_summaries %>% arrange(desc(n_authors)) %>%
    select(package, n_authors, n_imports, n_imported_by) %>% head(10)
package_summaries %>% arrange(desc(n_imports)) %>%
    select(package, n_authors, n_imports, n_imported_by) %>% head(10)
```

The sub-network of my packages
```{r}
my_packages <- subset(package_network$nodes, grepl("Ioannis Kosmidis", Author))$Package
package_summaries %>% filter(package %in% my_packages) %>%
    select(package, n_authors, n_imports, n_imported_by, eigen_centrality) %>% head(10)
visualize.cranly_network(package_network, package = my_packages)
```

# CRAN collaboration network

Build CRAN collaboration network
```{r}
author_network <- build_cranly_network(object = packages_db, perspective = "author")
```

Compute statistics for the collaboration network
```{r}
author_summaries <- summary(author_network)
```

Top 10 collaborators according to various network statistics
```{r}
library("dplyr")
author_summaries %>% arrange(desc(n_packages)) %>%
    select(author, n_packages) %>% head(10)
author_summaries %>% arrange(desc(page_rank)) %>%
    select(author, n_packages, page_rank) %>% head(10)
author_summaries %>% arrange(desc(betweenness)) %>%
    select(author, n_packages, betweenness) %>% head(10)
author_summaries %>% arrange(desc(eigen_centrality)) %>%
    select(author, n_packages, eigen_centrality) %>% head(10)
```

R Core's collaboration sub-network
```{r}
visualize.cranly_network(author_network, author = "R Core")
```
Brian Ripley's collaboration sub-network
```{r}
visualize.cranly_network(author_network, author = "Brian Ripley")
```
My collaboration sub-network (much much smaller that the above, of co)
```{r}
visualize.cranly_network(author_network, author = "Ioannis Kosmidis")
```

# A more focused analysis of the collaboration network

Let's focus on the sub-network with all authors having authored or contributing to more than 5 packages



<!-- Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format: -->

<!-- - Never uses retina figures -->
<!-- - Has a smaller default figure size -->
<!-- - Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style -->

<!-- ## Vignette Info -->

<!-- Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette. -->

<!-- ## Styles -->

<!-- The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows: -->

<!--     output:  -->
<!--       rmarkdown::html_vignette: -->
<!--         css: mystyles.css -->

<!-- ## Figures -->

<!-- The figure sizes have been customised so that you can easily put two images side-by-side.  -->

<!-- ```{r, fig.show='hold'} -->
<!-- plot(1:10) -->
<!-- plot(10:1) -->
<!-- ``` -->

<!-- You can enable figure captions by `fig_caption: yes` in YAML: -->

<!--     output: -->
<!--       rmarkdown::html_vignette: -->
<!--         fig_caption: yes -->

<!-- Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**. -->

<!-- ## More Examples -->

<!-- You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`. -->

<!-- ```{r, echo=FALSE, results='asis'} -->
<!-- knitr::kable(head(mtcars, 10)) -->
<!-- ``` -->

<!-- Also a quote using `>`: -->

<!-- > "He who gives up [code] safety for [code] speed deserves neither." -->
<!-- ([via](https://twitter.com/hadleywickham/status/504368538874703872)) -->
